---
title: "esm206_lab6"
author: "Russell Fager"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)


```

## Example of a rank-based test

we'll make our own samples using a pseudorandom generator

```{r}
set.seed(1414)

gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)

gp_2 <- sample.int(30, size = 15, replace = TRUE)

```

```{r}
hist(gp_1)

hist(gp_2)
```
Try a t.test

```{r}
t.test(gp_1, gp_2)
```

If these samples were drawn from populations with the same mean, the probability of taking 2 random samples with means *at least as different* as the sample means we found by random chance is 19.8%

retain (fail to reject) the null hypothesis. >> there is no signigicant difference in means between group 1 and group 2.

Cannot reject the null hypothesis (usually p-valuye must be less than 5% to reject).

Now lets compare this outcome to a rank based test.

## Mann Whitney U unpaired rank based test

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu
```

If these samples were drawn from populations with the same ranks (medians), the probability of finding two samples with ranks *at least as different* as those in our sample is 28%.

There is no significant difference in ranks (often you'll see medians instead of ranks) between group 1 and 2

retain (fail to reject) the null hypothesis. 


