---
title: "esm206_lab6"
author: "Russell Fager"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)


```

## Example of a rank-based test

we'll make our own samples using a pseudorandom generator

```{r}
set.seed(1414)

gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)

gp_2 <- sample.int(30, size = 15, replace = TRUE)

```

```{r}
hist(gp_1)

hist(gp_2)
```
Try a t.test

```{r}
t.test(gp_1, gp_2)
```

If these samples were drawn from populations with the same mean, the probability of taking 2 random samples with means *at least as different* as the sample means we found by random chance is 19.8%

retain (fail to reject) the null hypothesis. >> there is no signigicant difference in means between group 1 and group 2.

Cannot reject the null hypothesis (usually p-valuye must be less than 5% to reject).

Now lets compare this outcome to a rank based test.

## Mann Whitney U unpaired rank based test

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu
```

If these samples were drawn from populations with the same ranks (medians), the probability of finding two samples with ranks *at least as different* as those in our sample is 28%.

There is no significant difference in ranks (often you'll see medians instead of ranks) between group 1 and 2

retain (fail to reject) the null hypothesis. 

## Linear regression 

simple linear regression (single dependent variable, a single independent variable)

```{r}
ggplot(data = penguins, aes (x = flipper_length_mm, y = body_mass_g))+
  geom_point() +
  geom_smooth(method = "lm")

#if you add include = FALSE in the code chunk above (r) it wont show up in your final output.
```

Find a linear regression model using ordinary least squares describing the relationship between flipper length and body mass for these penguins.

3 piece:
- what type of model
- what is the relationship to model (DV ~ IV(s))
- where is the date that's used to create this model?

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

penguin_lm
```

On average I expect a 1 mm unit increase to be associated with a 49.69 increase in grams. 

what about the -5780.83? That's the body mass of a penguin if the flipper length was zero. Because this makes no sense you have to factor in the range. 

```{r}
penguin_lm_tidy <- broom::tidy(penguin_lm)
penguin_lm_tidy

# can use tidy or glance for different table summaries.

broom:: glance(penguin_lm)
```

How can I include my model equation in a report?

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

```{r}
plot(penguin_lm)
```

